---
title: "NRES 746 Snow-Climate Trends"
output: html_notebook
---
This code runs the mann-kendall test for multiple variables. The goal is to create one slope metric for each climate and snow variable so that we can run a random forest analysis on these data. 

```{r}
# load necessary libraries
library(raster)
library(rgdal)
library(parallel)
library(data.table)
library(gtools)
library(spatialEco)
library(EnvStats)
library(Kendall)
library(remotes)
library(snow)
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(rasterVis)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

The raster.kendall function will return errors because of the ci.slope input. I've copied the function from github below and specified ci.slope = FALSE. Otherwise, it is the same function as raster.kendall()
```{r}
raster_kendall <- function(x, intercept = FALSE, p.value = FALSE, z.value = FALSE,   
                           confidence = FALSE, tau = FALSE, ...) {
  if(!any(which(utils::installed.packages()[,1] %in% "EnvStats")))
    stop("please install EnvStats package before running this function")
  if(!any(class(x)[1] %in% c("RasterBrick","RasterStack"))) 
    stop("x is not a raster stack or brick object")
  if(confidence) {confidence = c(TRUE,TRUE)} else {confidence = c(FALSE,FALSE)}
    n <- c("intercept", "p.value", "z.value", "LCI", "UCI", "tau")	
	n <- n[which(c(intercept, p.value, z.value,confidence, tau))]	
    if( raster::nlayers(x) < 5) stop("Too few layers (n<5) to calculate a trend")
  trend.slope <- function(y, tau.pass = tau, p.value.pass = p.value,  
                          confidence.pass = confidence[1], z.value.pass = z.value,
                          intercept.pass = intercept) {
    fit <- suppressWarnings( EnvStats::kendallTrendTest(y ~ 1, ci.slope = FALSE) )
      fit.results <- fit$estimate[2]
        if(p.value.pass == TRUE) { fit.results <- c(fit.results, fit$p.value) } 
		  if(z.value.pass == TRUE) { fit.results <- c(fit.results, fit$statistic) } 
  	        if(confidence.pass == TRUE) { 
		      ci <- unlist(fit$interval["limits"])
		        if( length(ci) == 2) { 
		          fit.results <- c(fit.results, ci)
                } else {
                  fit.results <- c(fit.results, c(NA,NA))
                }			  
		    }
        if(intercept.pass == TRUE) { fit.results <- c(fit.results, fit$estimate[3]) }  
		  if(tau.pass == TRUE) { fit.results <- c(fit.results, fit$estimate[1]) }  
    return( fit.results )
  }
  k <- raster::overlay(x, fun=trend.slope, ...)
    names(k) <- c("slope", n)
  return( k )
}
```


```{r}
# create raster stacks
# get filenames and create a raster stack 
filenames <- list.files("wy_swe_max", full.names = T)
swe_stack <- stack(filenames)
# writeRaster(tmin_mam_stack, "tmin_MAM_stack.tif", overwrite = TRUE)
# get filenames and create a raster stack 
filenames <- list.files("tmin_MAM", full.names = T)
tmin_mam_stack <- stack(filenames)
# writeRaster(tmin_mam_stack, "tmin_MAM_stack.tif", overwrite = TRUE)

filenames <- list.files("tmin_DJF", full.names = T)
tmin_djf_stack <- stack(filenames)
# writeRaster(tmin_djf_stack, "tmin_DJF_stack.tif", overwrite = TRUE)

filenames <- list.files("precip_MAM", full.names = T)
precip_mam_stack <- stack(filenames)
# writeRaster(precip_mam_stack, "precip_MAM_stack.tif", overwrite = TRUE)

filenames <- list.files("precip_DJF", full.names = T)
precip_djf_stack <- stack(filenames)
# writeRaster(precip_djf_stack, "precip_DJF_stack.tif", overwrite = TRUE)
```


This takes a long time, so we parallelized the process. 
```{r}
# beginCluster(n=21) # n = number of cores
# 
# system.time(tmin_mam_mk <- raster::clusterR(tmin_mam_stack, raster_kendall, args=list(p.value = TRUE)))
# 
# endCluster()
```

```{r}
# beginCluster(n=21) # n = number of cores
# system.time(tmin_djf_mk <- raster::clusterR(tmin_djf_stack, raster_kendall, args=list(p.value = TRUE)))
# endCluster()
```


```{r}
# beginCluster(n=21) # n = number of cores
# system.time(precip_mam_mk <- raster::clusterR(precip_mam_stack, raster_kendall, args=list(p.value = TRUE)))
# endCluster()
```


```{r}
# beginCluster(n=21) # n = number of cores
# system.time(precip_djf_mk <- raster::clusterR(precip_djf_stack, raster_kendall, args=list(p.value = TRUE)))
# endCluster()
```



```{r}
# # write to rasters
# max_p_value_full <-max_trends_full_test[[2]] # define
# max_slope_full <-max_trends_full_test[[1]]
# writeRaster(max_p_value_full,"maxSWE_pval_mk.tif")
# writeRaster(max_slope_full, "maxSWE_slope_mk.tif")

# # write to rasters
# max_p_value_full <-tmin_mam_mk[[2]] # define
# max_slope_full <-tmin_mam_mk[[1]]
# writeRaster(max_p_value_full,"tmin_mam_pval_mk.tif", overwrite = TRUE)
# writeRaster(max_slope_full, "tmin_mam_slope_mk.tif", overwrite = TRUE)


# # write to rasters
# max_p_value_full <-tmin_djf_mk[[2]] # define
# max_slope_full <-tmin_djf_mk[[1]]
# writeRaster(max_p_value_full,"tmin_djf_pval_mk.tif", overwrite = TRUE)
# writeRaster(max_slope_full, "tmin_djf_slope_mk.tif", overwrite = TRUE)


# # write to rasters
# max_p_value_full <-precip_mam_mk[[2]] # define
# max_slope_full <-precip_mam_mk[[1]]
# writeRaster(max_p_value_full,"precip_mam_pval_mk.tif")
# writeRaster(max_slope_full, "precip_mam_mk.tif")
# 
# 
# # write to rasters
# max_p_value_full <-precip_djf_mk[[2]] # define
# max_slope_full <-precip_djf_mk[[1]]
# writeRaster(max_p_value_full,"precip_djf_pval_mk.tif")
# writeRaster(max_slope_full, "precip_djf_slope_mk.tif")


# # write to rasters
# max_p_value_full <-max_trends_full_test[[2]] # define
# max_slope_full <-max_trends_full_test[[1]]
# writeRaster(max_p_value_full,"maxSWE_pval_mk.tif")
# writeRaster(max_slope_full, "maxSWE_slope_mk.tif")
```

Now we have to factorize the data

```{r}
# load the rasters
swe_max_mk_slope <- raster("maxSWE_slope_mk.tif")
tmin_mam_mk_slope <- raster("tmin_mam_slope_mk.tif")
tmin_djf_mk_slope <- raster("tmin_djf_slope_mk.tif")
precip_mam_mk_slope <- raster("precip_mam_slope_mk.tif")
precip_djf_mk_slope <- raster("precip_djf_slope_mk.tif")

swe_max_mk_pval <- raster("maxSWE_pval_mk.tif")
tmin_mam_mk_pval <- raster("tmin_mam_pval_mk.tif")
tmin_djf_mk_pval <- raster("tmin_djf_pval_mk.tif")
precip_mam_mk_pval <- raster("precip_mam_pval_mk.tif")
precip_djf_mk_pval <- raster("precip_djf_pval_mk.tif")
```


```{r}
# save only significan values
swe_max_slope_sig <- swe_max_mk_slope
swe_max_slope_sig[swe_max_mk_pval$maxSWE_pval_mk > 0.05] <- NA
writeRaster(swe_max_slope_sig,"maxSWE_SigSlope_mk.tif", overwrite = T)

# save only significan values
tmin_mam_slope_sig <- tmin_mam_mk_slope
tmin_mam_slope_sig[tmin_mam_mk_pval$tmin_mam_pval_mk > 0.05] <- NA
# writeRaster(tmin_mam_slope_sig,"tmin_mam_SigSlope_mk.tif", overwrite = T)

# save only significan values
tmin_djf_slope_sig <- tmin_djf_mk_slope
tmin_djf_slope_sig[tmin_djf_mk_pval$tmin_djf_pval_mk > 0.05] <- NA
#writeRaster(tmin_djf_slope_sig,"tmin_djf_SigSlope_mk.tif", overwrite = T)


# save only significan values
precip_mam_slope_sig <- precip_mam_mk_slope
precip_mam_slope_sig[precip_mam_mk_pval$precip_mam_pval_mk > 0.05] <- NA
#writeRaster(precip_mam_slope_sig,"precip_mam_SigSlope_mk.tif")

# save only significan values
precip_djf_slope_sig <- precip_djf_mk_slope
precip_djf_slope_sig[precip_djf_mk_pval$precip_djf_pval_mk > 0.05] <- NA
#writeRaster(precip_djf_slope_sig,"precip_djf_SigSlope_mk.tif")

```
```{r}
# save swe variable as factor
swe_max_mk_slope_factorized_sig <- swe_max_slope_sig
swe_max_mk_slope_factorized_sig[swe_max_mk_slope_factorized_sig>0] <- 2
swe_max_mk_slope_factorized_sig[swe_max_mk_slope_factorized_sig<0] <- 1
swe_max_mk_slope_factorized_sig[swe_max_mk_slope_factorized_sig == 0] <- 0
swe_max_mk_slope_factorized_sig <- writeRaster(swe_max_mk_slope_factorized_sig, 'swe_max_mk_SigSlope_factorized.tif', datatype='INT2U', overwrite=TRUE)
```


Plot data and save to raster

```{r}
# jpeg(
#     filename="swe_mk.jpeg",
#     width=12,
#     height=10,
#     units="in",
#     res=500)
p1 <- levelplot(swe_max_mk_slope[[1]], names.attr="Slope", par.settings=RdBuTheme(),main="SWE Mann-Kendall Slope", at=seq(-0.02, 0.02, length.out=10))
p2 <- levelplot(swe_max_mk_pval[[1]], names.attr="Pvalue", par.settings=RdBuTheme(),main="SWE Mann-Kendall P-Value")
grid.arrange(p1, p2, ncol=2)
#dev.off()
```



```{r}
swe_mk_df <- as.data.frame(swe_max_mk_slope, xy=TRUE)
```


```{r}
cellStats(swe_max_mk_slope, stat = "sd", na.rm = TRUE)
```



```{r}
sig_p <- swe_max_mk_pval
sig_p[sig_p > 0.05] <- NA
sig_p[sig_p <= 0.5] <- 1

jpeg(filename="swe_change.jpg", width=10, height=8, units="in", res=500)
par(mfrow=c(1,2))
my.palette1 <- brewer.pal(n = 11, name = "RdBu")
p1 <- plot(swe_max_mk_slope, col = my.palette1, zlim = c(-0.02,0.02), main = "Change in Max SWE (cm/year)")
my.palette2 <- brewer.pal(n = 9, name = "Greys")
p2 <- plot(sig_p, col = my.palette2, main = "P-Values < 0.05 Max SWE")
dev.off()
```


```{r}
# specify only significant values
sig_p <- tmin_mam_mk_pval
sig_p[sig_p > 0.05] <- NA
sig_p[sig_p <= 0.5] <- 1

jpeg(filename="tmin_change.jpg", width=10, height=8, units="in", res=500)
par(mfrow=c(1,2), mar = c(5, 4, 4, 5) + 0.1)
my.palette1 <- brewer.pal(n = 11, name = "BuRd")
p1 <- plot(tmin_mam_mk_slope, col = my.palette1, main = "Change in Tmin MAM (mm/year)")
my.palette2 <- brewer.pal(n = 11, name = "RdBu")
p2 <- plot(tmin_djf_mk_slope, col = my.palette1, main = "Change in Tmin DJF (mm/year)")
dev.off()
```




```{r}
sig_p <- precip_djf_mk_pval
sig_p[sig_p > 0.05] <- NA
sig_p[sig_p <= 0.5] <- 1

jpeg(filename="precip_djf_change.jpg", width=10, height=8, units="in", res=500)
par(mfrow=c(1,2))
my.palette1 <- brewer.pal(n = 11, name = "RdBu")
p1 <- plot(precip_djf_mk_slope, col = my.palette1, zlim = c(-20,20), main = "Change in Precip DJF (mm/year)")
my.palette2 <- brewer.pal(n = 9, name = "Greys")
p2 <- plot(sig_p, col = my.palette2, main = "P-Values < 0.05 Precip DJF")
dev.off()
```



```{r}
sig_p <- precip_mam_mk_pval
sig_p[sig_p > 0.05] <- NA
sig_p[sig_p <= 0.5] <- 1

jpeg(filename="precip_change.jpg", width=10, height=8, units="in", res=500)
par(mfrow=c(1,2))
my.palette1 <- brewer.pal(n = 11, name = "RdBu")
p1 <- plot(precip_mam_mk_slope, col = my.palette1, zlim = c(-10,10), main = "Change in Precip MAM (mm/year)")
my.palette1 <- brewer.pal(n = 11, name = "RdBu")
p1 <- plot(precip_djf_mk_slope, col = my.palette1, zlim = c(-20,20), main = "Change in Precip DJF (mm/year)")
dev.off()
```

